Ternary plots for `hyper2` and `hyper3` samples

### 2024-10-18 inst/ternaryplot_hyper2.Rmd
use bookdown so figures can be numbered

---
title: "Ternary plots for `hyper2` and `hyper3` samples"
author: "Robin Hankin"
output: html_document
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

Using the `TernaryPlot()` and `TernaryPoints()` functions is straightforward:
```{r plotdata}
jj <- pnames(chess)
TernaryPlot(atip=jj[1],btip=jj[2],ctip=jj[3])
TernaryPoints(xx,cex=0.5)
TernaryPoints(chess_maxp,pch=16,col="red")
```

Above we see black circles being samples from the posterior and the
red point is the maximum likelihood estimate for the three players'
strengths.  We can adapt this technique to higher-dimensional `hyper2`
objects.

@@

```{r, iconsmatplot}
matplot(x,type="l",lty=1)
legend("topright",col=1:6,lty=1,legend=pnames(icons))
```

```{r, selectjustbigones}
wanted <- c(1,3) # NB,PB
xx <- cbind(x[,wanted],fillup=rowSums(x[,-wanted]))
TernaryPlot(atip="NB",btip="PB",ctip="other")
TernaryPoints(xx,cex=0.5)
```

@@

```{r ternaryrrace3}
TernaryPlot(atip="a", btip="b", ctip="c")
TernaryPoints(x, pch=16, cex=0.5)
TernaryPoints(c(0.3,0.5,0.2),pch=16,col="red")
```

Above, we see the ensemble of estimates of the players' Plackett-Luce
strengths as black dots, and the true value of the strengths
(viz. `c(a=0.3,b=0.5,c=0.2)`) in red.  The estimates cluster around the true value.  It might be interesting to assess whether the estimates 
estimates are Dirichlet and, if so, what parameters the distribution
possesses.

>>>>

---
title: "Ternary plots for `hyper2` and `hyper3` samples"
author: "Robin Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

Using the `TernaryPlot()` and `TernaryPoints()` functions is straightforward:
```{r plotdata, fig.cap="Ternary plot of estimated Plackett-Luce strengths of three chess players, using synthetic data"}
jj <- pnames(chess)
TernaryPlot(atip=jj[1],btip=jj[2],ctip=jj[3])
TernaryPoints(xx,cex=0.5)
TernaryPoints(chess_maxp,pch=16,col="red")
```

In figure \@ref(fig:plotdata) we see black circles being samples from
the posterior and the red point is the maximum likelihood estimate for
the three players' strengths.  We can adapt this technique to
higher-dimensional `hyper2` objects.

@@

```{r iconsmatplot, fig.cap="Metropolis-Hastings random sampling of the `icons` dataset: the six Bradley-Terry strengths of each icon"}
matplot(x,type="l",lty=1)
legend("topright",col=1:6,lty=1,legend=pnames(icons))
```

```{r selectjustbigones, fig.cap='Ternary plot of the icons dataset, showing Norfolk Broads ("NB", top), Polar Bears ("PB", lower right), and everything else ("other", lower left).  Small black dots show MLEs of synthetic (Metropolis-Hastings) datasets; large red dot shows MLE for the original dataset'}
wanted <- c(1,3) # NB,PB
xx <- cbind(x[, wanted], fillup=rowSums(x[, -wanted]))
TernaryPlot(atip="NB", btip="PB", ctip="other")
TernaryPoints(xx, cex=0.5, pch=16)
jj <- icons_maxp[wanted]
TernaryPoints(c(jj,1-sum(jj)), pch=16, col="red")
```

Figures \@ref(fig:iconsmatplot) and \@ref(fig:selectjustbigones) show
different aspects of the BT strengths of the six icons of the `icons`
dataset.

@@

```{r ternaryrrace3, fig.cap="Maximum likelihood estimates of synthetic datasets generated by the `rrace3()` random race function"}
TernaryPlot(atip="a", btip="b", ctip="c")
TernaryPoints(x, pch=16, cex=0.5)
TernaryPoints(c(0.3,0.5,0.2),pch=16,col="red")
```

Figure \@ref(fig:ternaryrrace3) shows an ensemble of estimates of the players' Plackett-Luce
strengths as black dots, and the true value of the strengths
(viz. `c(a=0.3,b=0.5,c=0.2)`) in red.  The estimates cluster around the true value.  It might be interesting to assess whether the estimates 
estimates are Dirichlet and, if so, what parameters the distribution
possesses.
